Performance of prediction models for opportunities for improvment in trauma care among women compared to men and across patient cohorts

### 2023-09-23 project-plan.Rmd
Comment on project plan

Traumatic injuries<!-- Call it just trauma--> are a significant public
health concern worldwide, accounting for a substantial portion of
morbidity and mortality rates, being the cause of more than 5 million
fatalities each year [@WHO2014]<!-- This is an old reference, most
current figure is 4.3 million. Use:
@article{injuries2020,
  title =	 {Injuries—Level 1 cause},
  volume =	 {396},
  ISSN =	 {0140-6736},
  journal =	 {The Lancet},
  publisher =	 {Elsevier BV},
  author =	 {{GBD 2019 Diseases and Injuries Collaborators}},
  year =	 {2020},
  month =	 {Oct},
}
-->.  Trauma, and care of patients
with traumatic injuries, have long been a significant target point for
healthcare improvement programs [@WHO2009].  Multidisciplinary
mortality and morbidity conferences form the cornerstone of
initiatives dedicated to enhancing the quality of trauma care and,
consequently, improving patient outcomes [@Santana2014].

@@

Audit filters are applied to determine possible events in patient care
that are associated with unfavourable outcomes. The occurrence of such
an event triggers a review process, and when appropriate leads to the
implementation of corrective measures to address systematic errors and
feedback to individual practitioners [@McDermott_1994]. The use of
audit filters has its drawbacks, being associated with high rates of
false positives. [Saknar tydlig källa här]<!-- Consider changing the
order of these paragraphs, so that you introduce audit filters before
mortality and morbidity conferences, considering that audit filters
are used to select patients for those conferences-->

@@

# Methods
<!-- I would find this section easier to read if you followed some
reporting guideline, like TRIPOD--> This is a registry-based study,
encompassing all trauma patients enlisted in both the Karolinska
University Hospital trauma registry and the trauma care quality
database spanning from 2012 to 2022. The Karolinska University
Hospital in Solna is equivalent to a level 1 trauma centre and manages
approximately 1,500 acute trauma patients each
year[@Ghorbani2018][Swetrau, rapport]. The trauma registry comprises
all patients admitted to the hospital with trauma team activation,
irrespective of their Injury Severity Score (ISS), as well as patients
with an ISS exceeding 9. Included in the registry is data on vital
signs, injuries, interventions, times, and patient demographics.

@@

All patients reviewed for possible opportunities for improvement between
the years 2013 and 2021 are included in the study, with the exception of
patients under the age of 15 due to differing clinical pathways in this
subgroup.

@@

Currently, there is no need to develop a thorough backup plan since all
the necessary data for the statistical analysis has been collected and
is accessible.

>>>>

Traumatic injuries<!-- Call it just trauma--> <!-- KS: when removing the word injuries it can be good to define "trauma" to differ from psychological trauma,
A good citation is:
@book{gerdin_risk_2015,
    location = {Stockholm},
    title = {The risk of dying: predicting trauma mortality in urban Indian
             hospitals},
    isbn = {978-91-7676-157-1},
    shorttitle = {The risk of dying},
    publisher = {Karolinska Institutet},
    author = {Gerdin, Martin},
    date = {2015},
    note = {{OCLC}: 942942790},
}
A good read if you have time ;), PDF URL: https://openarchive.ki.se/xmlui/bitstream/handle/10616/44832/Thesis_Martin_Gerdin.pdf?sequence=3&isAllowed=y
--> 
are a significant public
health concern worldwide, accounting for a substantial portion of
morbidity and mortality rates, being the cause of more than 5 million
fatalities each year [@WHO2014]<!-- This is an old reference, most
current figure is 4.3 million. Use:
@article{injuries2020,
  title =	 {Injuries—Level 1 cause},
  volume =	 {396},
  ISSN =	 {0140-6736},
  journal =	 {The Lancet},
  publisher =	 {Elsevier BV},
  author =	 {{GBD 2019 Diseases and Injuries Collaborators}},
  year =	 {2020},
  month =	 {Oct},
}
-->.  Trauma, and care of patients
with traumatic injuries, have long been a significant target point for
healthcare improvement programs [@WHO2009].  Multidisciplinary
mortality and morbidity <!-- KS: The order "morbidity and mortality" is mainly used in the literature. I also received a comment on this in my final version. --> conferences form the cornerstone of
initiatives dedicated to enhancing the quality of trauma care and,
consequently, improving patient outcomes [@Santana2014].

@@

Audit filters are applied to determine possible events in patient care
that are associated with unfavourable outcomes. The occurrence of such
an event triggers a review process, and when appropriate leads to the
implementation of corrective measures to address systematic errors and
feedback to individual practitioners [@McDermott_1994]. The use of
audit filters has its drawbacks, being associated with high rates of
false positives. [Saknar tydlig källa här]<!-- Consider changing the
order of these paragraphs, so that you introduce audit filters before
mortality and morbidity conferences, considering that audit filters
are used to select patients for those conferences-->
<!-- KS: you will probably need to explain exactly what audit fitlers are. That they are a system that flags abnormal clinical parameters as possible OFI. These flag cases are then reviewed by two nurses and then a conference -->

@@

# Methods
<!-- I would find this section easier to read if you followed some
reporting guideline, like TRIPOD--> <!-- KS: prefer "We conducted a registry[...]" -->This is a registry-based study,
encompassing all trauma patients enlisted in both the Karolinska
University Hospital trauma registry and the trauma care quality
database spanning from 2012 to 2022. The Karolinska University
Hospital in Solna is equivalent to a level 1 trauma centre and manages
approximately 1,500 acute trauma patients each
year[@Ghorbani2018][Swetrau, rapport]. The trauma registry comprises
all patients admitted to the hospital with trauma team activation,
irrespective of their Injury Severity Score (ISS), as well as patients
with an ISS exceeding 9. Included in the registry is data on vital
signs, injuries, interventions, times, and patient demographics.

@@

All patients reviewed for possible opportunities for improvement <!-- KS: use OFI consistently after you have abbreviated it --> between
the years 2013 and 2021 are included in the study, with the exception of
patients under the age of 15 due to differing clinical pathways in this
subgroup.

@@

<!-- KS: Gianluigi will most likely not like this explaination so if you have time just come up with a random backup project -->
Currently, there is no need to develop a thorough backup plan since all
the necessary data for the statistical analysis has been collected and
is accessible.
